Contextual Identity: Towards Usable Privacy

### 2013-03-03 paper.tex
still rewriting last section

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information.  The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}. Helen
Nissenbaum has long argued that privacy violations come not from the simple
sharing of personal information, but rather from sharing that information in a
way that violates social norms, or in the wrong context~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth
who are at higher risk for bullying and suicide than non-LGTB youth, and much
of that support is only accessible through the Internet in socially
conservative areas~\cite{hrc}. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertantly outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, because Queer Chorus is a
public group Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Many authors have discovered methods to link different contextual identities to
the same person. Narayanan and Shmatikov present a re-identification technique
to merge anonymized social graphs from different networks and prove that
user identifiers from Netflix and IMDB and Flickr and Twitter can be
linked~\cite{narayanan1,narayanan2}. Lindamood et al.~and Mislove et al.~show
how to infer previously undisclosed information from public social networking
data~\cite{lindamood,mislove}.

@@

\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private messages,
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

@@

Even worse, many of these providers now serve as login platforms for external
sites, spreading these unsatisfiable policies outside the service providers
themselves.  Facebook Connect is a login platform that allows third-party
websites to authenticate users using their Facebook identity~\cite{fb_connect}.
Because it is against the terms of service to have multiple Facebook accounts,
using Facebook Connect has the unwanted side-effect of linking multiple
contextual identities.

@@

\subsection{How do users think about identity?}
We hypothesize many users are not consciously aware of having multiple
contextual identities. In order to develop tools to help users, we need to
understand users' mental models of identity and how information is shared on
the Internet. For computer scientists, authentication is tied to identity, and
even in the absence of authentication, using the same device over time
implicitly creates an identity through tracking techniques and local
information. However, these points are far from obvious to a typical user,
especially for users who don't distinguish the web application from the
browser, from the operating system, or from the device.

@@

\begin{itemize}
\item How is identity represented online?
\item Is your identity tied to the device you are using?
\item If you check your mail, read news, log on to a social network at a
computer at the public library, does that affect your identity?
\item What does it mean to share devices? If you lend your tablet to your
sister, is she representing herself, or you?
\item If you register for an account on a service, do you expect that to have
an influence on your online identity? What if you never use that account again?
\item If you visit a website, do you expect that to have an influence on your
browsing experience or online identity? If so, for how long?
\item Which parts of your online identity do you expect to be visible to
your housemates, friends, relatives, or employers?
\item Which parts of your identity do you want to share with whom?
\item Which parts of your identity do you want to keep secret from whom?
\item How does your identity change over time?
\end{itemize}

danah boyd has done excellent work surveying youth and social networks, and Pew
Internet is a good source for self-reported phone surveys on privacy and social
media, but there is otherwise a dearth of research in these
areas~\cite{boyd,pew1,pew2,pew3}.

@@

\paragraph{Multiple accounts}
The long-time existence of support for multiple accounts in email clients
suggests that many users have multiple email identifiers, which could be
considered as a proxy for identity if tied to service accounts. Similarly, data
suggests that a large minority of Twitter users have multiple Twitter accounts,
and the plethora of cross-posting or syndication software from one social
network to another suggests that many users take advantage of multiple social
networks~\cite{twitter}.
\begin{comment}
That is not to say that all users have the same
tolerance or even need for multiple explicit identities: in fact many users may
be exclusively task-oriented and use the Internet with specific tasks in mind
rather than for social interaction or exploratory surfing\footnote{Even these
users are subject to building implicit identities through the use of tracking
techniques, however.}.
\end{comment}

\paragraph{Auditing}
Wang et.~al suggest that common techniques for handling regret online include
manual deletion of regrettable posts, self-censoring or delaying
posts that they predict might bring regret~\cite{wang}.

\paragraph{Private browsing mode}
Private browsing mode exists in all major browsers, but are not a complete
solution~\cite{ABBJ10}. There is no standard behavior for private browsing
mode, and interaction with extensions, treatment of cookies, history, and
bookmarks upon entering and exiting this mode are different and serve different
use cases. For example, entering private browsing in Firefox does not clear
cookies, history, or other session data; session data is cleared on exiting
private browsing mode. In contrast, incognito mode in Chrome makes previous
session state unavailable and clears new session state on exit; this mode
may be preferable as a ``Guest mode'' where the device owner does not want the
guest to either view or pollute existing state.

\paragraph{Profile management}
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so unusable
that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}

\paragraph{Cookie blockers}
Disconnect and ShareMeNot are browser extensions to disallow interacting with
service providers like Google and Facebook unless users explicitly choose to
interact with that site~\cite{disconnect,franzi}.

@@

\paragraph{Mitigating explicit linkage}
Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\paragraph{Controlling Information Flow}
Do Not Track is an HTTP request header that signals to the website that the
user does not want to be tracked~\cite{dnt}. Although Do Not Track is
unenforceable from the client side, service providers who ignore signal will
lose reputation by acting against their users' explicit wishes.

\paragraph{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience. A browser is in an ideal position to intermediate social network
posts, aggregate them locally, and present them to the user when they desire to
audit their digital footprint.  For example, a user could see all of the
comments and posts they made in the last week and redact content that is
sensitive or too negative, a frequent cause of regret~\cite{wang}.  Although
this approach does not help when the service provider decides to change the
audience on the user's behalf, auditing gives the user a good opportunity to
review posts that might cause regrets. In the case of high-volume users, such a
tool could incorporate sentiment analysis to aid the user in finding
particularly problematic content.

\paragraph{Implicit identity generation}
Profiles can be generated by user-supplied application-level data, but also
implicitly through browsing via tracking techniques.  Collusion is a tool for
visualizing cookies~\cite{collusion}, in particular third-party cookies set by
tracking sites which are typically ad networks.  Once the user knows which
sites are tracking them, they can choose to block cookies from those sites.

>>>>

The desire for privacy balances the desire for spontaneous, positive human
interaction, which often requires sharing personal information. The act of
sharing alone does not negate expectation of privacy~\cite{boyd2}.
%The desire for spontaneous, positive human interaction often requires sharing
%personal information. Sharing information can lead to loss of privacy, though
%it does not negate expectation of privacy~\cite{boyd2}. 
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}.

Lesbian, gay, bisexual, and transgendered youth are at risk for contextual
identity violations.  Access to support is crucial for LGTB-identified youth,
who are at higher risk for bullying and suicide than non-LGTB youth~\cite{hrc}.
Much of that support is only accessible through the Internet in socially
conservative areas. In 2011 Bobbi Duncan and Taylor McCormick, two
students at the University of Texas at Austin, were inadvertantly outed when
their choir director added them to a Facebook group for Queer
Chorus~\cite{fowler}. Even though both students had blocked their parents from
seeing any personal posts revealing their sexuality, because Queer Chorus is a
public group Facebook published notifications to Duncan's and McCormick's
friends, including their parents. Duncan later attempted suicide~\cite{duncan}.

@@

Many authors have discovered methods to link different contextual identities to
an individual. Narayanan and Shmatikov present a re-identification technique
to merge anonymized social graphs from different networks and prove that
user identifiers from Netflix and IMDB and Flickr and Twitter can be
linked~\cite{narayanan1,narayanan2}. Lindamood et al.~and Mislove et al.~show
how to infer previously undisclosed information from public social networking
data~\cite{lindamood,mislove}.

@@

\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more
to the user's activity stream.  Because of poor opt-out and lack of visibility
into what was being published, user outcry was
immediate~\cite{mccarthy,nakashima}.  In December 2007, Google Reader exposed
RSS feeds of user-marked news stories to the user's Google Talk contacts, which
includes everyone with whom a user has chatted whether they be co-workers,
supervisors, or friends. Although this feed was always public, prior to this
launch it was not discoverable, leading to a bad experience for many
users~\cite{helft}. In February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts, which users used to treat as private messages
before the advent of ``Like'' and comment buttons, to the attention of an
audience that the user never anticipated~\cite{timeline}.

@@

Even worse, many of these providers now serve as login platforms for external
sites. For example, Facebook Connect allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}.  Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities across multiple sites.

@@

\subsection{How do users think about identity?}
We hypothesize many users are not consciously aware of having multiple
contextual identities. In order to develop tools to help users, we need to
understand users' mental models of identity and how information is shared on
the Internet. For computer scientists, authentication is tied to identity, and
even in the absence of authentication, using the same device over time
implicitly creates an identity through tracking techniques and local
information. However, these points are far from obvious to a typical user,
especially for users who don't distinguish the web application from the
browser, from the operating system, or from the device\footnote{What is a browser? \url{https://www.youtube.com/watch?v=o4MwTvtyrUQ}} \footnote{\url{http://input.mozilla.org/en-US/?q=facebook&product=firefox}}.

@@

\begin{itemize}
\item How is identity represented online?
\item What do you need to represent your identity?
\item Is your identity tied to your device?
\item If you check your mail, read news, log on to a social network at a
computer at the public library, does that affect your identity?
\item What does it mean to share devices? If you lend your tablet to your
sister, is she representing herself, or you?
\item If you register for an account on a service, do you expect that to have
an influence on your online identity? What if you never use that account again?
\item If you visit a website, do you expect that to have an influence on your
browsing experience or online identity? If so, for how long?
\item Which parts of your online identity do you expect to be visible to
your housemates, friends, relatives, or employers?
\item Which parts of your online identity do you expect to be visible to
websites you use? What about websites you don't use?
\item Which parts of your identity do you want to share with whom?
\item Which parts of your identity do you want to keep secret from whom?
\item How does your identity change over time?
\end{itemize}

danah boyd's work surveying youth and social networks is excellent, and Pew
Internet is a good source for self-reported phone surveys on privacy and social
media, but there is otherwise a dearth of research in these
areas~\cite{boyd,pew1,pew2,pew3}.
% digital natives is related but not exactly

@@

\subsubsection{Separating identities}

The following examples illustrate how users keep their identities separate.

\paragraph{Multiple accounts}
The long-time existence of support for multiple accounts in email clients
suggests that many users have multiple email identifiers, which could be
considered as a proxy for identity if tied to service accounts. Similarly, data
suggests that a large minority of Twitter users have multiple Twitter accounts,
and the plethora of cross-posting or syndication software from one social
network to another suggests that many users take advantage of multiple social
networks~\cite{twitter}.

\paragraph{Multiple browsers}

Using multiple browsers is a useful technique for managing multiple accounts.
Some services support multiple accounts but not multiple login (e.g.,
Twitter). For services that do (e.g., many Google services), the user interface
may be so confusing that it leads to errors, so the best recommendation may be
to use multiple browsers~\cite{yegge}.\footnote{Google recommends multiple
browsers as an alternative to multiple login:
\url{https://support.google.com/accounts/bin/answer.py?hl=en&answer=179235}}.
Additionally, for users who want to separate work and
personal browsing, using multiple browsers is the easiest solution.

\paragraph{Multiple devices}

Many users have multiple devices and many use them for different purposes.
Sometimes the difference in purpose is due to the nature of the device (e.g.,
GPS and mapping software are probably more often used in mobile devices) and
some may be due to policy (e.g., personal browsing is limited to a personal
laptop). Using multiple devices implicitly creates multiple browser states and
thus multiple identities.  However, these identities may still be linked
through authentication data or other techniques~\cite{aol}.

\paragraph{Multiple profiles}

Using multiple browsers forces separation between cookies, history, builtin
password managers, and other local storage. However, several browsers support
separation of profile data natively. Firefox has a Profile Manager, but it is
so unusable that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.} Google Chrome
supports multiple profiles\footnote{Manage multiple users on Chrome:
\url{https://support.google.com/chrome/bin/answer.py?hl=en&answer=2364824}}. In
both cases, the threat model does not include users with malicious intent in
the same household.

\paragraph{Private browsing mode}
Private browsing mode exists in all major browsers, but are not a complete
solution~\cite{ABBJ10}. There is no standard behavior for private browsing
mode: interaction with extensions, treatment of cookies, history, and
bookmarks upon entering and exiting are different across browsers.
For example, entering private browsing in Firefox does not clear
cookies, history, or other session data; session data is cleared on exiting
private browsing mode. In contrast, incognito mode in Chrome makes previous
session state unavailable and clears new session state on exit; this mode
may be preferable as a ``Guest mode'' where the device owner does not want the
guest to either view or pollute existing state. L

\paragraph{Cookie blockers}
Disconnect and ShareMeNot are browser extensions to disallow interacting with
service providers like Google and Facebook unless users explicitly choose to
interact with that site~\cite{disconnect,franzi}.  Collusion is a tool for
visualizing cookies~\cite{collusion}, in particular third-party cookies set by
tracking sites which are typically ad networks.  Once the user knows which
sites are tracking them, they can choose to block cookies from those sites.

\paragraph{Controlling Information Flow}
Do Not Track is an HTTP request header that signals to the website that the
user does not want to be tracked~\cite{dnt}. Although Do Not Track is
unenforceable from the client side, service providers who ignore signal
lose reputation by acting against their users' explicit wishes.

\paragraph{Service settings} Service configurations such as Facebook friends,
Google circles and privacy settings are intended for users to self-manage
privacy and identity. However, shifting privacy settings and complex,
interacting features make these configurations unpredictable for many users.

\subsubsection{Linking identities}
\paragraph{Linking between profiles}

-link from fb profile to blog
-link between services

\paragraph{Synchronizing browser data}

- Chrome sync
- picl
- xmarks

@@

\paragraph{Mitigating explicit linkage}
Federated login services can prevent linking contextual identities to the same
user identifier. For example, Persona is the suite of tools that implements the
BrowserID protocol, and it supports multiple email
identifiers~\cite{browserid}.  If the type of service is known, Persona could
prevent the user from associating the same email identifier to radically
different contextual identities, such as ones for dating and professional use.

\paragraph{Auditing}
(needs rewrite)
A major cause of user embarrassment is exposing information to the wrong
audience. A browser is in an ideal position to intermediate social network
posts, aggregate them locally, and present them to the user when they desire to
audit their digital footprint.  For example, a user could see all of the
comments and posts they made in the last week and redact content that is
sensitive or too negative, a frequent cause of regret~\cite{wang}.  Although
this approach does not help when the service provider decides to change the
audience on the user's behalf, auditing gives the user a good opportunity to
review posts that might cause regrets. In the case of high-volume users, such a
tool could incorporate sentiment analysis to aid the user in finding
particularly problematic content.
